The Increased Effect of Elections and Changing Prime Ministers on Topics Discussed in the Australian Federal Parliament between 1901 and 2018

### 2019-10-31 outputs/docs/paper/paper_MA_structure_edit.Rmd
rest of methods

There are 14,680 days of publicly available Hansard records across the two chambers of the Australian Federal Parliament for which we have PDFs. Further summary statistics for this are provided in Appendix \@ref(hansardsummarystatistics). In general, the frequency of sitting days based on information from the Hansard dataset compared to the actual number of sitting days is comparable, almost complete coverage of our dataset. XX MAYBE pull out a few more summary stats here XX

@@

The output of interest from the topic modelling stage, is the proportion of each topic appearing in each document. The aim of the analysis stage of the modelling process is to analyse how the distribution of topics changes in relation to different types of events. But with many topics for each of the roughly 14,680 chamber-sitting-days the data are still too noisy to easily visualise changes around events.

One option for relating the topic distributions to events would be to use the Structural Topic Model (STM) of @RobertsStewartAiroldi2016. The distinguishing aspect of the STM is that it considers more than just a document's content when constructing topics. For example, we may believe a document's author, or the time at which it was written, or, in the case of our paper, the prime minister or election period, may affect the topics within that document. The STM allows this additional information, or metadata, to affect the construction of topics, though influencing either topical prevalence or topical content. The assumption that there is some document generation process is the same as in LDA, it is just that this process now includes metadata.

@@

To overcome these challenges, we formalise a statistical framework that allows us to systematically identify significant changes in topic distributions over time. Specifically, we use the estimated topic distributions from the CTM described in the previous section as an input into a Bayesian hierarchical Dirichlet regression framework, which relates the proportions of each topic to underlying time trends, changes in prime minister and elections. This set-up also allows us to identify 'outlying' topic distributions and relate these to other events.

Define $\theta_{cdp}$ to be the proportion of topic $p$ on day $d$ in chamber $c$ (where chamber refers to either the House of Representatives or the Senate). Note that the $\theta_{c,d,1:P}$ for $p = 1,2,\dots, P = 19$ are equal to the estimated values of $\theta_{cd}$ from the CTM after we group them into the Comparative Agendas Project topics. We assume that the majority of variation in topics by day is across sitting periods, $s$, where a sitting period is defined as any group of days that are less than one week apart. Using this definition, there are 822 sitting periods over the period 1901 to 2018 inclusive. Appendix \@ref(hansardsummarystatistics) contains more information about the sitting patterns over the course of the year, which have changed considerably since Federation.

@@

The goal of the model is to relate these proportions to the current prime minister $g$ and election period $e$, assuming that an 'election effect' would be at its peak in the days close to the election, then decay over the period. The mean parameters $\mu_{c,s,p}$ are modelled on the log scale as:
$$
\log \mu_{c,s,p} = \alpha_{g[c,s],p} + \beta_{e[c,s],p} \cdot (N_{s[e]} - s) + \delta_{c,s,p}
$$
where: $\alpha_{g[c,s],p}$ is the mean effect for prime ministers $g$ (which covers sitting period $s$ in chamber $c$) and topic $p$; $\beta_{e[c,s],d,p}$ is the effect of election $e$ (which occurs in sitting period $s$, in chamber $c$) for topic $p$; $N_s$ is the total number of sitting periods in election period $e$; and $\delta_{c,s,p}$ is a random, or levels, effect for each chamber, sitting period and topic.

The term for the prime minister, $\alpha_{g[c,s],p}$, assumes there is some underlying mean effect of each prime minister on the topic distribution that is common across chambers. We place uninformative priors on each of these parameters:
$$
\alpha_{g[c,s],p} \sim \mbox{Normal}(0, 100).
$$

@@

Finally, the sitting-period-specific random effect $\delta_{c,s,p}$ allows for chamber- and sitting-period specific deviations in the topic distributions. This allows us to identify periods where there may be differences in the topics discussed across the two chambers. It also allows us to identify large deviations away from the expected distribution, thus helping to identify the effect of other, non-prime-minister and non-election events. In addition, this set up also partially pools effects across sitting periods. The $\delta_{c,s,p}$ values are modelled as:
$$
\delta_{c,s,p} \sim \mbox{Normal}(0, \sigma_{g[c,s],p}^2).
$$

The variance parameters $\sigma_{g[c,s],p}^2$ give an indication of the how the variation in topics is changing over government periods. If the estimates of the variance are larger, then there is more variation in the topics discussed within a government period. Non-informative priors are placed on the variance parameters:
$$
\sigma_{g[c,s],p} \sim \mbox{Uniform}(0,3).
$$

>>>>

There are 14,680 days of publicly available Hansard records across the two chambers of the Australian Federal Parliament for which we have PDFs. Further summary statistics for this are provided in Appendix \@ref(hansardsummarystatistics). In general, the frequency of sitting days based on information from the Hansard dataset compared to the actual number of sitting days is comparable, almost complete coverage of our dataset. XX MAYBE PULL OUT MORE SUMMARY STATS IN PARTICULAR HIGHLIGHT THE TWO CHAMBERS XX

@@

The output of interest from the topic modelling stage is the proportion of each topic appearing in each document. The aim of the analysis stage of the modelling process is to analyse how the distribution of topics changes in relation to different types of events. But with many topics for each of the roughly 14,680 chamber-sitting-days spanning 118 years, the data are still too noisy to easily visualise changes around events.

One option for relating the topic distributions to events would be to use the Structural Topic Model (STM) of @RobertsStewartAiroldi2016. The distinguishing aspect of the STM is that it considers more than just a document's content when constructing topics. For example, we may believe a document's author, or, in the case of our paper, the prime minister or election period, may affect the topics within that document. The STM allows this additional information, or metadata, to affect the construction of topics, though influencing either topical prevalence or topical content. The assumption that there is some document generation process is the same as in LDA, it is just that this process now includes metadata.

@@

To overcome these challenges, we formalise a statistical framework that allows us to systematically identify significant changes in topic distributions over time. Specifically, we use the estimated topic distributions from the previous step as an input into a Bayesian hierarchical Dirichlet regression framework, which relates the proportions of each topic to underlying time trends, changes in prime minister and elections.

### Model set-up

Define $\theta_{c,d,p}$ to be the proportion of speech that refers to topic $p$ on day $d$ in chamber $c$ (where chamber refers to either the House of Representatives or the Senate). Note that the $\theta_{c,d,1:P}$ for $p = 1,2,\dots, P = 19$ are equal to the estimated values of $\theta_{cd}$ from the CTM after we group them into the Comparative Agendas Project topics. We assume that the majority of variation in topics by day $d$ is across sitting periods, $s$, where a sitting period is defined as any group of days that are less than one week apart. Using this definition, there are 822 sitting periods over the period 1901 to 2018 inclusive. Appendix \@ref(hansardsummarystatistics) contains more information about the sitting patterns over the course of the year, which have changed considerably since Federation.

@@

The goal of the model is to relate these proportions to the current prime minister $g$ and election period $e$, assuming that an 'election effect' would be at its peak in the days close to the election, then decay over the period. The mean parameters $\mu_{c,s,p}$ are modelled on the log scale as:

$$
\log \mu_{c,s,p} = \alpha_{g[c,s],p} + \beta_{e[c,s],p} \cdot (N_{s[e]} - s) + \delta_{c,s,p}
$$
where: $\alpha_{g[c,s],p}$ is the mean effect for prime ministers $g$ (which covers sitting period $s$ in chamber $c$) and topic $p$; $\beta_{e[c,s],d,p}$ is the effect of election $e$ (which occurs in sitting period $s$, in chamber $c$) for topic $p$; $N_s$ is the total number of sitting periods in election period $e$; and $\delta_{c,s,p}$ is a random, or levels, effect for each chamber, sitting period and topic.

The term for the prime minister, $\alpha_{g[c,s],p}$, assumes there is some underlying mean effect of each prime minister on the topic distribution that is common across chambers. Non-informative priors are placed on $\alpha_{g[c,s],p}$:

$$
\alpha_{g[c,s],p} \sim \mbox{Normal}(0, 100).
$$

@@

Finally, the sitting-period-specific random effect $\delta_{c,s,p}$ allows for chamber- and sitting-period specific deviations in the topic distributions. This allows us to identify periods where there may be differences in the topics discussed across the two chambers. It also allows us to identify large deviations away from the expected distribution, thus helping to identify the effect of other, non-prime-minister and non-election events. In addition, this set up also partially pools effects across sitting periods. The $\delta_{c,s,p}$ values are modelled heirarchically as:

$$
\delta_{c,s,p} \sim \mbox{Normal}(\mu_{c,p}, \sigma_{g[c,s],p}^2).
$$
This structure assumes that the sitting period effect is a draw from a distribution with a mean that is common to that particular chamber and topic, with some associated variance. This allows for the sitting-period-specific random effect $\delta_{c,s,p}$ to be partially informed by other sitting periods for that chamber and topic combination.

The variance parameters $\sigma_{g[c,s],p}^2$ give an indication of the how the variation in topics is changing over government periods. If the estimates of the variance are larger, then there is more variation in the topics discussed within a government period. Non-informative priors are placed on the variance parameters:

$$
\sigma_{g[c,s],p} \sim \mbox{Uniform}(0,3).
$$
